---
title: "Antares new features v9.2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Antares new features v9.2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(antaresEditObject)
```

This article will present the new features in line with **Antares v9.2** (cf  [Antares Simulator](https://antares-simulator.readthedocs.io/en/latest/user-guide/04-migration-guides/#v920))

## New features

**Short Term Storages**  :

  - Parameter `group` is now dynamic and have no restriction for user  


## Create new study v9.2

```{r init}
dir_path <- tempdir()
suppressWarnings(
  createStudy(path = dir_path, 
            study_name = "test920", 
            antares_version = "9.2")
)

```

For Antares Simulator, format version is now "9.2". For packages, we make the adjustment and we keep "920". 

**Check version of my current study** : 

```{r}
current_study_opts <- simOptions()
current_study_opts$antaresVersion
```


## Create area 

We just need create *areas* to create *st-storages*. 
```{r areas}
createArea(name = "fr")
createArea(name = "it")
```

## Short term storage

We can create new clusters, st-storage (from v8.6), with function `createClusterST()`. You can see function documentation with `?createClusterST`.  

By default you can call function only with two parameters (`area`, `cluster_name`). 

Clusters are created with default properties and time series.

### Dynamic group names

Default `group` is still "Other1" and now you can **create/edit** your own group name (*only for version study >= 9.2*).

```{r st-storage/group}
# creation
createClusterST(area = "fr", 
                cluster_name = "test_storage", 
                group = "my_own_group")

createClusterST(area = "it", 
                cluster_name = "test_storage", 
                group = "my_own_group_again")

# edit group of existing st-storage cluster
editClusterST(area = "fr", 
              cluster_name = "test_storage", 
              group = "my_own_group_Pondage")

# read cluster properties
tab <- readClusterSTDesc()
rmarkdown::paged_table(tab)
```
### New properties

you can create or edit new clusters with new properties (see doc `?createClusterST`). 

```{r default values}
# new properties (default values)
rmarkdown::paged_table(as.data.frame(storage_values_default(), check.names = FALSE))
```

```{r creation/properties}
# creation
my_parameters <- storage_values_default()
my_parameters$efficiencywithdrawal <- 0.5
my_parameters$`penalize-variation-injection` <- TRUE
my_parameters$`penalize-variation-withdrawal` <- TRUE

createClusterST(area = "fr", 
                cluster_name = "test_storage", 
                group = "new_properties", 
                storage_parameters = my_parameters, 
                overwrite = TRUE)

createClusterST(area = "it", 
                cluster_name = "test_storage", 
                group = "new_properties", 
                storage_parameters = my_parameters,
                overwrite = TRUE)

# read cluster properties 
tab <- readClusterSTDesc()
rmarkdown::paged_table(tab)
```

```{r edit/properties}
# edit properties of existing st-storage cluster
my_parameters$efficiencywithdrawal <- 0.9
my_parameters$`penalize-variation-injection` <- FALSE
my_parameters$`penalize-variation-withdrawal` <- FALSE

editClusterST(area = "fr", 
              cluster_name = "test_storage",
              storage_parameters = my_parameters)

# read cluster properties 
tab <- readClusterSTDesc()
rmarkdown::paged_table(tab)
```

### New optional **time series**

We have five new *.txt* files containing one series of dimension ${N=8760, P=1}$ :  

  - cost-injection.txt  
  - cost-withdrawal.txt 
  - cost-level.txt  
  - cost-variation-injection.txt  
  - cost-variation-withdrawal.txt
  

```{r create/ts}
# creation
ratio_value <- matrix(0.7, 8760)
  
# default properties with new optional TS
createClusterST(area = "fr", 
                cluster_name = "good_ts_value", 
                cost_injection = ratio_value, 
                cost_withdrawal = ratio_value, 
                cost_level = ratio_value, 
                cost_variation_injection = ratio_value,
                cost_variation_withdrawal = ratio_value)

# read cluster TS values 
tab <- readInputTS(st_storage = "all", 
                   showProgress = FALSE)
rmarkdown::paged_table(head(tab))
```

```{r edit/ts}
# edit TS values of existing st-storage cluster
new_ratio_value <- matrix(0.85, 8760)

# edit everything or anyone you want 
editClusterST(area = "fr",
              cluster_name = "good_ts_value",
              cost_injection = new_ratio_value, 
              cost_withdrawal = new_ratio_value)

# read cluster TS values 
tab <- readInputTS(st_storage = "all", 
                   showProgress = FALSE)
rmarkdown::paged_table(head(tab))
```

### Remove clusters

Nothing has changed to remove clusters.

```{r remove}
# read cluster names
levels(readClusterSTDesc()$cluster)

# remove a cluster
removeClusterST(area = "fr", 
                cluster_name = "good_ts_value")

# read cluster 
tab <- readClusterSTDesc()
rmarkdown::paged_table(tab)
```


## Generaldata

The “generaldata.ini” settings file (*settings/generaldata.ini*) contains several sections.

Antares Simulator v9.2 deletes some parameters: 

  - *adequacy patch/enable-first-step*
  - *adequacy patch/set-to-null-ntc-between-physical-out-for-first-step*
  - *other preferences/initial-reservoir-levels*

A message is displayed and parameters are set to `NULL` for the `[adequacy patch]` section.

For the `[other preferences]` section, the `initial-reservoir-levels` parameter is not explicitly used by a dedicated function.

```{r generaldata}
# user messages
updateAdequacySettings(
    set_to_null_ntc_between_physical_out_for_first_step = FALSE)
updateAdequacySettings(enable_first_step = FALSE)
```

```{r delete study, include=FALSE}
# Delete study
unlink(current_study_opts$studyPath, 
       recursive = TRUE)
# clean global options
options(antares = NULL)
```





