---
title: "Antares new features v9.2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Antares new features v9.2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(antaresEditObject)
```

This article will present the new features in line with **Antares v9.2** (cf  [Antares Simulator](https://antares-simulator.readthedocs.io/en/latest/user-guide/04-migration-guides/#v920))

## New features

**Short Term Storages**  :

  - Parameter `group` is now dynamic and have no restriction for user  


## Create new study v9.2

```{r init}
dir_path <- tempdir()
suppressWarnings(
  createStudy(path = dir_path, 
            study_name = "test920", 
            antares_version = "9.2")
)

```

For Antares Simulator, format version is now "9.2". For packages, we make the adjustment and we keep "920". 

**Check version of my current study** : 

```{r}
current_study_opts <- simOptions()
current_study_opts$antaresVersion
```


## Create area 

We just need create *areas* to create *st-storages*. 
```{r areas}
createArea(name = "fr")
createArea(name = "it")
```

## Short term storage

We can create new clusters, st-storage (from v8.6), with function `createClusterST()`. You can see function documentation with `?createClusterST`.  

By default you can call function only with two parameters (`area`, `cluster_name`). 

### No restriction on group names

Default `group` is still "Other1" and now you can **create/edit** your own group name (*only for version study >= 9.2*).

```{r st-storage}
# creation
createClusterST(area = "fr", 
                cluster_name = "test_storage", 
                group = "my_own_group")

createClusterST(area = "it", 
                cluster_name = "test_storage", 
                group = "my_own_group_again")

# edit group of existing st-storage cluster
editClusterST(area = "fr", 
              cluster_name = "test_storage", 
              group = "my_own_group_Pondage")

# read cluster properties 
tab <- readClusterSTDesc()
rmarkdown::paged_table(tab)
```

```{r delete study, include=FALSE}
# Delete study
unlink(current_study_opts$studyPath, 
       recursive = TRUE)
# clean global options
options(antares = NULL)
```