---
title: "Variant management with API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Variant management with API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

```{r setup}
library(antaresEditObject)
```

The API behind [Antares Web](https://antares-web.readthedocs.io/en/latest/) comes with a Variant Manager allowing to edit a study. Functions from {antaresEditObject} can be used send commands to the API or generate commands to be sent to the API.


## Path to simulation and variant creation

First we need to declare which study we are going to use:

```{r set-simulation-path, eval=FALSE}
antaresRead::setSimulationPathAPI(
  host = "http://localhost:8080",
  study_id = "70a08fae-da67-444a-b2ed-df4c0f956a31", 
  token = NULL, 
  simulation = "input"
)
```

Then we can create a new variant from our study or use one created through the web interface:

```{r create-variant, eval=FALSE}
# Create new variant
createVariant("variant-1")

# use an existing one
useVariant("variant-2")
```


A third option is to mock the web server behavior, it can be useful if we are offline or if we just want to generate API commands to be use afterwards:

```{r mock-api}
mockSimulationAPI()
```


## API editing mode

{antaresEditObject} allow to use two modes to use the API:

* **async**: record all API calls, but nothing is sent to the server, commands can be retrieved after calling {antaresEditObject}'s functions
* **sync**: send query to the API each time a function is used (commands are also internally recorded if needed).

```{r set-api-mode, eval=FALSE}
setAPImode("async")
# or
setAPImode("sync")
```

Default is to used **async** mode. When using `mockSimulationAPI()` only **async** mode is available.



## Get / export API commands

API commands generated after calling functions like `createArea()`, `createLink()`, ... can be retrieved at all time with:

```{r get-api-commands, eval=FALSE}
getAPIcommands()
```

Last command generated can be viewed with:

```{r get-api-commands-last, eval=FALSE}
getAPIcommands(last = TRUE)
# or use a numeric to get the last N commands
getAPIcommands(last = 3)
```

You can also filter type of commands with:

```{r get-api-commands-actions, eval=FALSE}
getAPIcommands(actions = "create_area")
```

Export commands with:

```{r write-api-commands, eval=FALSE}
writeAPIcommands("path/to/commands.json")
```



## Usage example

Below are listed all functions from {antaresEditObject} that can be used with the API. These functions will include the following badge in their documentation:

```{r, echo=FALSE}
knitr::include_graphics("figures/badge_api_ok.svg")
```


### Create area


```{r, eval=FALSE}



# Set mode you want to use
# setAPImode("sync") # send all commands to API
# setAPImode("async") # just record all commands


# Create a New Variant
createVariant("variant-1")

# or use a pre-existing one:
# useVariant("variant-2")


# Create a new area
createArea(name = "earth")

# List commands recorded (one normally at this point)
getAPIcommands()

# Create another area with more options
createArea(
  name = "moon", 
  filtering = filteringOptions(filter_synthesis = c("hourly", "daily"))
)

# List commands
getAPIcommands()

```

