---
title: "Variant management with API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Variant management with API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

```{r setup}
 # CRAN limite CPU usage
data.table::setDTthreads(2)
library(antaresEditObject)
```

The API behind [Antares Web](https://antares-web.readthedocs.io/en/latest/) comes with a Variant Manager allowing to edit a study. Functions from {antaresEditObject} can be used send commands to the API or generate commands to be sent to the API.


## Path to simulation and variant creation

First we need to declare which study we are going to use:

```{r set-simulation-path, eval=FALSE}
antaresRead::setSimulationPathAPI(
  host = "http://localhost:8080",
  study_id = "70a08fae-da67-444a-b2ed-df4c0f956a31", 
  token = NULL, 
  simulation = "input"
)
```

Then we can create a new variant from our study or use one created through the web interface:

```{r create-variant, eval=FALSE}
# Create new variant
createVariant("variant-1")

# use an existing one
useVariant("variant-2")
```


A third option is to mock the web server behavior, it can be useful if we are offline or if we just want to generate API commands to be use afterwards:

```{r mock-api}
mockSimulationAPI()
```


## API editing mode

{antaresEditObject} allow to use two modes to use the API:

* **async**: record all API calls, but nothing is sent to the server, commands can be retrieved after calling {antaresEditObject}'s functions
* **sync**: send query to the API each time a function is used (commands are also internally recorded if needed).

```{r set-api-mode, eval=FALSE}
setAPImode("async")
# or
setAPImode("sync")
```

Default is to used **async** mode. When using `mockSimulationAPI()` only **async** mode is available.



## Get / export variant management commands

Variant commands generated after calling functions like `createArea()`, `createLink()`, ... can be retrieved at all time with:

```{r get-api-commands, eval=FALSE}
getVariantCommands()
```

Last command generated can be viewed with:

```{r get-api-commands-last, eval=FALSE}
getVariantCommands(last = TRUE)
# or use a numeric to get the last N commands
getVariantCommands(last = 3)
```

You can also filter type of commands with:

```{r get-api-commands-actions, eval=FALSE}
getVariantCommands(actions = "create_area")
```

Export commands with:

```{r write-api-commands, eval=FALSE}
writeVariantCommands("path/to/commands.json")
```



## Usage example

Below are listed all functions from {antaresEditObject} that can be used with the API. These functions will include the following badge in their documentation:

![](https://img.shields.io/badge/Antares%20API-OK-green)


### Create an area

Create a new area:

```{r create_area}
createArea(name = "area01")
createArea(name = "area02")
createArea(name = "area03")
getVariantCommands()
```

Create a second area with some default parameters:

```{r create_area2}
createArea(
  name = "area04", 
  filtering = filteringOptions(filter_synthesis = c("hourly", "daily"))
)
getVariantCommands()
```


You can also edit an area or remove it:

```{r remove_area}
createArea(name = "area000")
# editArea(name = "area000", ...)
removeArea(name = "area000")
getVariantCommands(last = TRUE)
```


### Create a link

Create a new link between two areas like this:

```{r create_link}
createLink(from = "area01", to = "area02")
createLink(from = "area01", to = "area03")
getVariantCommands(last = 2)
```

Edit an existing link with:

```{r edit_link}
editLink(
  from = "area01", 
  to = "area02",
  dataLink = matrix(data = c(rep(9, 8760*2), rep(6, 8760*6)), ncol = 8)
)
getVariantCommands(last = 2)
```

Remove a link with:

```{r remove_link}
removeLink(from = "area01", to = "area03")
getVariantCommands(last = TRUE)
```


### Create a cluster

Create a new cluster with:

```{r create_cluster}
createCluster(
  area = "area01", 
  cluster_name = "clus01"
)
getVariantCommands(last = TRUE)
```

With more parameters:

```{r create_cluster2}
createCluster(
  area = "area01", 
  cluster_name = "clus02",
  unitcount = 1L,
  marginal_cost = 50,
  ts_interpretation = "production-factor",
  group = "Nuclear",
  add_prefix = FALSE,
  prepro_data = matrix(
    data = c(rep(9, times = 365 * 2),
             rep(7, times = 365 * 4)), 
    ncol = 6
  ),
  prepro_modulation = matrix(
    data = c(rep(8, times = 365 * 24 * 3),
             rep(6, times = 365 * 24 * 1)),
    ncol = 4
  ),
  time_series = matrix(
    data = c(rep(22, times = 365 * 24 * 8),
             rep(44, times = 365 * 24 * 4)),
    ncol = 12
  )
)
getVariantCommands(last = 2)
```


Edit a cluster with:

```{r edit_cluster, eval = FALSE}
createCluster(
  area = "area02", 
  cluster_name = "clus02"
)
editCluster(
  area = "area02", 
  cluster_name = "clus02", 
  unitcount = 5L
)
getVariantCommands(last = TRUE)
```


Remove a cluster with (`removeCluster()` has been updated with endpoint and no longer uses an api command ):

```{r remove_cluster, eval=FALSE}
createCluster(
  area = "area02", 
  cluster_name = "clus000"
)
removeCluster(
  area = "area02", 
  cluster_name = "clus000"
)
getVariantCommands(last = TRUE)
```





### Create a binding constraint

Create a new binding constraint with:

```{r create_binding_constraint}
createBindingConstraint(
  name = "myconstraint",
  values = NULL,
  enabled = FALSE,
  timeStep = "hourly",
  operator = "both",
  coefficients = c("area01%area02" = 1)
)
getVariantCommands(last = TRUE)
```

You can edit a binding constraint with `editBindingConstraint()` and remove one with `removeBindingConstraint()`.


### Update settings

Update general settings:

```{r update_general_settings}
updateGeneralSettings(mode = "Adequacy", generate = c("thermal", "hydro"))
getVariantCommands(last = 2)
```


Update input settings:

```{r update_input_settings}
updateInputSettings(import = c("hydro", "thermal"))
getVariantCommands(last = TRUE)
```


Update optimization settings:

```{r update_optimization_settings}
updateOptimizationSettings(
  simplex.range = "week",
  power.fluctuations = "minimize ramping"
)
getVariantCommands(last = 2)
```


Update output settings:

```{r update_output_settings}
updateOutputSettings(
  synthesis = TRUE,
  storenewset = FALSE,
  archives = c("load", "wind")
)
getVariantCommands(last = TRUE)
```


### Scenario builder

Read data from scenario builder (here's it's empty since we're mocking the API):

```{r read_scenario_builder}
readScenarioBuilder()
```

Create new data to use as scenario builder, since we're mocking the API parameters must be set explicitly, otherwise thez are retrieved from the study.

```{r scenario_builder}
my_scenario <- scenarioBuilder(n_scenario = 3, areas = c("area01", "area02"), n_mc = 10)
my_scenario
```

Then you can update the scenario builder itself:

```{r update_scenario_builder}
updateScenarioBuilder(ldata = my_scenario, series = "load")
getVariantCommands(last = TRUE)
```


### Other functions

Write input time series:

```{r write_input_ts}
writeInputTS("area01", type = "solar", data = matrix(rep(4, 8760*2), nrow = 8760))
getVariantCommands(last = TRUE)
```

Write water values:
```{r write_water_values}
writeWaterValues("area01", data = matrix(rep(0, 365*101), nrow = 365))
getVariantCommands(last = TRUE)
```





