---
title: "Renewables Energy Sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Renewables Energy Sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(antaresEditObject)
```


Antares v8.0.3 allow to work with renewable energy sources, implemented in Antares as a new functionnality: renewables clusters.
Here's a quick start guide to use those new features.

As always, setup the study you want to use:

```{r}
tmp <- tempfile()
createStudy(path = tmp, study_name = "res-study")
opts <- setSimulationPath(tmp)
```



## Activate RES

First you need to activate RES functionality, to do so you can use:

```{r}
activateRES()
```

This will set `renewable-generation-modelling` parameters to `clusters` and create appropriate file structure to create RES clusters.


You can also update `renewable-generation-modelling` parameters manually with:

```{r}
updateOptimizationSettings(renewable.generation.modelling = "clusters")
```


## Manipulate clusters

Like thermal clusters, there's equivalent functions to create, edit and remove RES clusters:

```{r}
# create an area to create a cluster in it
createArea(name = "area51")

# Create a renewable cluster
createClusterRES(
  area = "area51", 
  cluster_name = "ren01", 
  add_prefix = FALSE
)

# Use an other group and some parameters
createClusterRES(
  area = "area51", 
  cluster_name = "ren02", 
  group = "Wind Offshore",
  nominalcapacity = 123,
  ts_interpretation = "production-factor",
  add_prefix = FALSE
)
```


Edit an existing cluster with:

```{r}
editClusterRES(
  area = "area51", 
  cluster_name = "ren02", 
  group = "Solar Rooftop", # new group
  add_prefix = FALSE
)
```


And remove a cluster with:

```{r}
removeClusterRES(area = "area51", cluster_name = "ren01")
```



## Scenario builder

There's a new type of serie allowed in scenario builder for renewable clusters called `r`:

```{r}
readScenarioBuilder()
```

```r
$r
      [,1] [,2] [,3]
area51   1    2    3  
```

To update the scenario builder, you can use:

```r
# generate scenario
sbuilder <- scenarioBuilder(n_scenario = 51)

# Update with:
updateScenarioBuilder(ldata = sbuilder, series = "renewables")
# OR
updateScenarioBuilder(ldata = list(r = sbuilder))
```




